A game-theory modeling approach to utility and strength of interactions dynamics in biomedical research social networks

### 2017-01-10 casm.tex
Title page and casm version

\documentclass[11pt]{article}                     
\usepackage{graphicx}
\usepackage[section]{placeins}
\usepackage{color}
\usepackage{lineno}
\linenumbers
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{A game-theory modeling approach to utility
  and strength of interactions dynamics in biomedical research social networks}
%\title{Two games modeling approach to collaboration dynamics in biomedical
%research social networks}

%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}

%\subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Fitness \& trust in biomedical research networks}

%\titlerunning{Short form of title}        % if too long for running head

\author{J. Mario Siqueiros-Garc\'ia         \and      Rodrigo Garc\'ia-Herrera \and         Enrique Hern\'andez-Lemus \and
        Sergio A. Alcal\'a-Corona  %etc.
}

%\authorrunning{Short form of author list} % if too long for running head

%% \institute{J. Mario Siqueiros-Garc\'ia \at
%%   IIMAS-UNAM. Circuito Escolar 3000, Cd Universitaria, Coyoacán, Ciudad de México, D.F. \\
%%               Tel.: +123-45-678910\\
%%               Fax: +123-45-678910\\
%%               \email{jmario.siqueiros@iimas.unam.mx}           %  \\
%% %             \emph{Present address:} of F. Author  %  if needed
%%               \and
%%               Rodrigo Garc\'ia-Herrera,\\ Enrique Hern\'andez-Lemus, \\ Sergio A. Alcal\'a-Corona \at
%%               National Institute of Genomic Medicine\\
%%               Perif\'erico Sur 4809, Arenal Tepepan, Tlalpan, 14610, \\
%%               Ciudad de M\'exico, D.F., M\'exico  
%%            %% S. Author \at
%%            %%    second address
%% }

\date{\today}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}

Collaboration has become a cornerstone in biomedical research today.
In contrast to physics which has a long history and experience in
collaborative projects, biology is only recently becoming an evermore
collaborative discipline.  In this article we explore the effect of
collaboration on the distribution of players' access to resources from
other players in the network and the distribution of the strength of
interactions among them. Particularly, we implemented two games played
simultaneously: one for maximizing individual utility based on the
iterated Prisoner's Dilemma; the other, a coordination game for
maximizing the connection strength between players. We are interested
in how they affect each other in the context of a network of
scientific collaboration under the idea that while researchers are
interested in maximizing their individual utilities, they also know
that it is important to invest in building collaborative
relationships.  We tested our simulation on a biomedical research
community network of M\'exico and compared the results with an
Erd\"{o}s-Reny\'i, a Watts-Strogatz small-world and Barab\'asi-Albert
topologies. Different topologies display different utility and
interaction strength distributions. Moreover, the distribution of
utility and interaction strength in the researchers network is similar
to that of Barab\'asi-Albert and Watts-Strogatz topologies,
respectively. We believe that utility distribution in the researchers
network suggests that there are socio-cultural mechanisms governing
the network that produce an asymmetric distribution of resources. The
high distribution of strong interactions might reflect some sort of
subordination among researchers by which they are morally obliged to
cooperate by the same socio-cultural mechanisms. The range around the
threshold that regulates the decision to cooperate or defect according
to the agent's historical balance between utility and strength of
collaborative relationships and carrying capacity of the system is
small, suggesting that there is a region in which a phase transition
takes place from a population of cooperators to a population of
defectors. Simulations like this may help to develop science policies
to promote fair distribution of resources.

\end{abstract}

\section{Introduction}
\label{intro}

%% \textbf{Notes:}

%% \begin{itemize}

%% %% \item In this paper there are two games played simultaneously: One for
%% %% maximazing individual utility; the second for maximazing the connection
%% %% strength. We are interested in how they affect eachother in the context of a
%% %% network of collaboration.

%% \item We change utility for fitness. Utility seems a better concept in order to
%%   reflect players' preferences.

%% %% \item In the context of our paper, utility represents
%% %%   the preference for having access to the resources of others. Maximazing
%% %%   utility means that the player is increasing its access to the resources of
%% %%   other players.  Utility function is just the sum of the payoffs of playing
%% %%   with its neighbors. As for the payoff matrix, the best strategy for a player
%% %%   is to defect. Any other option will not be an optimal payoff. The opossing
%% %%   force is in the other game. The game for maximazing the strength of the
%% %%   connection among players. In this case defecting or cooperating? is the best
%% %%   strategy as it pays the most when coordinating regardeless they both cooperate
%% %%   or defect, this is because when both cooperate the interaction gets a positive
%% %%   payoff, when both defect, the interaction doesn't get affected; but if they
%% %%   anti-coordinate, then the interaction looses. In this sense, this might be a
%% %%   game were there are multiple (i.e., 2 points or strategies) equilibria. It
%% %%   seems that there is no single dominant strategy.}\\

%% \item Choosing a strategy depends on the balance $\eta'$ equation --look at page
%%   9. This
%%   equation states that the probability of an agent to cooperate increases as: a)
%%   its utility increases --as it moves away from 0, or; b) The strength of the
%%   connection with its peers increases --as it moves away from 0, or; c) both
%%   increases.

%%   $\eta_i = \frac{\langle f_i \rangle + \langle w_{ij} \rangle _j}{2}$\\

%% \item{En la pagina 9 decimos que ``In order for the agent to choose to cooperate
%%   or defect, a degree of willingness to cooperate is assigned to each agent.'' No estoy
%%   seguro que ``degree of willingness to cooperate'' sea la mejor manera de
%%   decirlo, mi duda está en la palabra ``degree'', por lo que cambi\'e la
%%   oraci\'on a: The probability for an agent to cooperate or defect depends on a
%% number that referes to a historical balance between average utility and the average strength of
%% the connections with its neigbors. }

%% \end{itemize}

Collaboration has become a cornerstone in biomedical research today.
In contrast to physics which has a long history and experience in
collaborative projects, biology is only recently becoming an evermore
collaborative discipline\cite{Vermeulen2013}. Biology has an
interesting record in such matters because scientific collaboration
means something different to different branches of biology: molecular
biology has traditionally been a research activity of small
laboratories\cite{KnorrCetina1999,Strasser2006}, whereas in natural
history there has been data and samples exchange since the $XVII^{th}$
century\cite{Muller2012,Strasser2012}. Despite the differences in
culture and practices, the Human Genome Project made collaboration a
central feature of biology.\\

@@

The analysis of cooperation in scientific research has been the subject of a
number of studies
\cite{Vermeulen2013,Newman2001,Newman2004,Elango2012,HernandezLemus2013,Strasser2006,Strasser2012}. This
is not surprising since cooperation and competition are quite important in
today's academic success. How does collaboration happen within a competitive
academic environment and what kind of payoff is present in these settings were
questions considered recently by Wardil and Hauert \cite{Wardil2015} in the
context of cooperation in multi-authored publications. Also, the role of game
theory over complex scientific information and collaboration networks has
attracted attention, mainly focusing on how long-term strategies may shape
different scenarios for Nash equilibria \cite{hanauske2012}. Prisoner's Dilemma
has been used in the study of impact factor and collaboration
\cite{Hara_etal_2002,Lieberman_etal_2005}. \\

Even with all these research efforts, cooperation in the context of
scientific collaboration is still loosely defined and the long term
dynamics of academic cooperation (and its consequences) are yet to be
fully elucidated. Furthermore, to our current knowledge, there has
been no use of game theory and complex network analysis for
understanding how the topology of scientific collaboration networks
affects access to resources among individuals present in the network
\footnote{For an account of scientific collaboration and definitions,
  please refer to \cite{Sonnenwald2007}.}. Our work aims to contribute
to our current understanding on the matter, specially when agents have
to maximize their access to resources while taking care of their collaboration links.\\

  %% nevertheless
  %% and to a certain degree, the idea behind the model is ethnographically
  %% inspired\footnote{For a mixed methodology approach to scientific
  %%   collaboration refer to Hara et al., 2002 \cite{Hara_etal_2002}.}. In
  %% previous work we observed that members of biomedical research groups interact
  %% strongly with each other but interactions between groups are only
  %% superficial. Our ethnographic work also allowed us to see how difficult is for
  %% Principal Investigators to collaborate with peer groups that are considered
  %% somehow equal in terms of different criteria such a productivity or power. On
  %% the other hand, collaboration may be easier with groups that are far beyond in
  %% some of these terms. This seems to be so because when dealing with a smaller
  %% group, the stronger one can establish conditions of collaboration; when the
  %% group is weaker, there might be a lot to win. For example, this latter
  %% situation can be seen when local groups collaborate with research groups form
  %% overseas, specially from the US.}\\

%% {\color{red}The question behind our model rests on three premises: Scientist
%%   have self-interests and may act seeking their own interests; Scientific
%%   collaboration needs researchers to cooperate and build relationships that can
%%   last; Scientific collaboration networks have a particular topology. Under such
%%   context, if the same amount of resources were given to every researcher at the
%%   same point in time, how would the access to resources be distributed, in the
%%   long run, in a population of interacting researchers that seek to improve
%%   their profits but at the same time have to invest in building trustworthy
%%   relationships with their colleagues?}
%%\\

  %% Seen from a different point of view, how would trust be
  %% distributed in a population of researchers that are mainly interested in
  %% improving their profits, even when acting selfishly always could have a
  %% negative impact on that primary objective?}\\

 In this article we explore the network effect on the distribution
   of players having access to certain amount of resources shared by other players in the
   network and the distribution of the strength of interactions among
   them. Particularly, we implemented two games played simultaneously: 
   one for maximizing individual utility based on the iterated Prisoner's
   Dilemma; the other, a coordination game for maximizing the connection strength between
   players. We are interested in how they affect each other in 
   the context of a network of scientific collaboration under the idea that while
   researchers are interested in maximizing their utilities, they also know that
   it is important to invest in building collaborative relationships. These two
   behaviors are explored in a biomedical research community of M\'exico.\\

In the context of our paper, utility represents
  access to resources shared by others.
  The value of the Utility function for a player is the sum of the payoffs of playing
  with its neighbors. The opposing force comes from the other concurrent game: players
  trying to maximize the strength of their interactions with other players. In the coordinating game
  the best strategy is to adopt the same strategy as the other player, as it pays the most
  regardless of cooperation or defection in the utility game. When
  both cooperate the interaction gets a positive payoff, when both defect, the
  interaction doesn't get affected; but if they anti-coordinate, then the
  interaction looses. Finally, cooperation is a central feature of scientific work.
  For our biomedical network, cooperation can be thought of
  as sharing resources such as time, students, equipment, even money. Examples of defection
  to a cooperator are ghost authorship or
  prestige authorship.\\

The manuscript is structured in five sections. First we describe
\textit{FOSISS}, the main program for grants destined to biomedical applied
research in M\'exico. This is the source of the
database from which we created the researchers collaboration network. Next we
describe our model
and the different network topologies on which we explored it.
We then present our results and discuss them. In the last section
we draw some final remarks and conclusions.

%Our main objective is to achieve a better understanding of the role of network
%topology on scientific collaborative social dynamics. The dynamics of the
%system are explored following the distribution of utility -as a proxy for
%resources- and trust as agents on the network play the Prisoner's Dilemma while
%constrained by trust with their neighbors.\\

%% In order to explore these questions we implemented a game theory simulation on a
%% real social network of biomedical researchers from M\'exico. Game theoretical
%% approaches to cooperation have – of course– been studied for decades. Taking
%% into account the nature of social interaction among players by means of an
%% underlying social network is also an explored topic. Some 10 years ago, Lo and
%% colleagues \cite{Lo2004} formulated a minority game model to connected
%% population within a network and studied the role of connectivity on such
%% games. Later on, Buesser and Tomassini \cite{Buesser2012} studied the dynamics
%% of cooperation for players located on a spatially embedded network in order to
%% categorize the effect that different interaction structures may have in
%% maintaining cooperation (or defection) domains.\\

%% The impact that network topology may have in the robustness of cooperation was
%% investigated in numerical experiments and simulations \cite{Ichinose2013}
%% revealing that clusters of cooperating agents become robust by increasing the
%% number of links to cooperators, with the possible implication that a large
%% number of individuals is required to maintain global cooperation patterns.\\

%% Regarding possible strategies to improve cooperation patterns, several mechanisms such as inequity aversion
%% \cite{Ahmed2014}, social imitation and strategic choice \cite{Vilone2014} have been explored. Inequity aversion is the
%% condition in which individuals care about payoff equality in the outcomes. A kind of moral memory in past outcomes
%% implies that inequity aversion promotes cooperation \cite{Ahmed2014}. Social imitation contrasts with strategic choices
%% in such a way that social influence models may impact on background rationality in decision making. This balance may be
%% behind moody conditional cooperation within a social environment \cite{Vilone2014}.\\

%% The analysis of cooperation in scientific research has been also the subject of a number of studies. This is not
%% surprising since cooperation and competitionare quite important in today’s academic success. How do collaboration 
%% happens within a competitive academic environment and what kind of payoff
%% is present in these settings were questions considered recently by Wardil and
%% Hauert \cite{Wardil2015} in the context of cooperation in multi-authored
%% publications. Also, the role of game theory over complex scientific information 
%% and collaboration networks has attracted attention, mainly focusing on how
%% long-term strategies may shape different scenarios for Nash equilibria \cite{hanauske2012}. \\

%% In spite of all these research efforts, cooperation in the context of scientific
%% collaboration is still loosely defined and the long time dynamics of academic
%% cooperation (and its consequences) are yet to be fully elucidated. Furhtermore,
%% to our current knowledge, we are not aware of any work on how the topology of
%% scientific collaboration networks affects the differences of access among
%% individuals to resources present on the network \footnote{For
%%   an account of scientific collaboration and definitions, please refer to
%%   \cite{Sonnenwald2007}.}. In this context our work aims to contribute to
%% further our current understanding on the matter.\\

%% As has been pointed by others, social research on scientific
%% collaboration is mostly about physics and space research\cite{Vermeulen2013}.
%% Collaboration in biology has been studied to a lesser degree
%% \cite{Newman2001,Newman2004,Elango2012,HernandezLemus2013,Strasser2006,Strasser2012}. The
%% use of game theory, and specifically the Prsioner's Dilemma to understand social
%% dynamics in scientific collaboration can be found at 
%% \cite{Hara_etal_2002,Lieberman_etal_2005}.\\

%% The manuscript is structured in five sections. The following describes \textit{FOSISS}, the main program for grants
%% destined for biomedical applied research in M\'exico. This description is important because that is the source of our
%% database for creating the researchers collaboration network. Next we describe our model, and then, in the following
%% section we introduce the different network topologies on which we explored the model. In the fourth section we present
%% our results and discuss them briefly. In the fifth section we draw some final remarks and conclusions.

\section{Biomedical research: CONACyT and FOSISS}
\label{sec:1}

@@

\emph{FOSISS} or Sectoral Fund for Health and Social Security Research
(\emph{Fondo Sectorial en Investigaci\'on en Salud y Seguridad Social}) is one of
such funds. FOSISS is constituted by CONACyT, SSA, IMSS and ISSSTE,\footnote{SSA
  is the acronym for Secretariat of Health \emph{Secretar\'ia de Salud}; IMSS is
  the acronym for Social Security Mexican Institute (\emph{Instituto Mexicano
    del Seguro Social}); ISSSTE stands for Institute for Social Security and
  Services for State Workers (\emph{Instituto de Seguridad y Servicios Sociales
    de los Trabajadores del Estado})} all of them being the major public health
providers and research institutions in the country. Every year CONACyT opens a
call for funds limited to a set of health research areas previously defined by a
group of experts. Such areas range from public health issues to chronic and
degenerative diseases.\\

Most applicants are public universities and research institutions, but
eligibility is open to public and private health research sectors.
From 2002 through 2013, there were 91 institutions funded that
comprised 4988 researchers.\\

@@

\section{Methodology}
\label{sec:2}
%% In this work we developed a model in order to have a better understanding of the
%% distribution of utility and trust in real networks of biomedical research
%% collaboration under certain social constraints  or topological properties.\\

Our model is based on the iterated version of the Prisoner's Dilemma (PD) and a
coordination game instantiated on
networks. Implementing games on networks is not new and it's an active area of
research aimed to understand the evolution of cooperation in networks populated
by selfish agents \cite{Szabo2007,Nowak1992,Nowak2006,Santos2005,Santos2006}. In many network models
on which some of game theory games are simulated, agents' decision to cooperate
or defect depend on a specific strategy, such as the well known
\textit{tit-for-tat} \cite{Axelrod2006,Nowak2011}. In some other cases, agents
can modify the weight of the interactions with their neighbors
\cite{Santos2006}. From a different perspective others have explored the effect
of different topologies on the emergence of cooperation
\cite{Santos2005,Hauert2004}. In our model, an agent's decision to cooperate or
defect depends on a balance between
utilities and the current strength of its collaboration relationships. Such
balance reflects the overall success or failure of its strategies. We
study the behavior of the system under different topologies, including a
real-world network.\\

@@

\subsection{Network initialization and agent state update}

All networks are initialized equally. The number of nodes for every network is 4122, the same as in the FOSISS network.
The same utility is given to every agent and all edges are assigned the same weight. In the case of the FOSISS
network, edge weight is given by the number of collaborations among researchers, utility remains the same for all nodes as in
the other networks.\\

The probability for an agent to cooperate or defect depends on a
number ($\eta$) that represents a historical balance between average utility and
the average strength of the connections with its neighbors. This is so
  because we assume that whatever the result in utility or strength of
  connection, as long as one of them increases, the player will be confident in
  the strategy followed so far.\\

@@

\subsection{Erd\"{o}s-Reny\'i}

>>>>

%% BioMed_Central_Tex_Template_v1.06
%%                                      %
%  bmc_article.tex            ver: 1.06 %
%                                       %

%%IMPORTANT: do not delete the first line of this template
%%It must be present to enable the BMC Submission system to
%%recognise this template!!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                     %%
%%  LaTeX template for BioMed Central  %%
%%     journal article submissions     %%
%%                                     %%
%%          <8 June 2012>              %%
%%                                     %%
%%                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% For instructions on how to fill out this Tex template           %%
%% document please refer to Readme.html and the instructions for   %%
%% authors page on the biomed central website                      %%
%% http://www.biomedcentral.com/info/authors/                      %%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%% BioMed Central currently use the MikTex distribution of         %%
%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
%% http://www.miktex.org                                           %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

%%% loading packages, author definitions

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
\documentclass{bmcart}

%%% Load packages
%\usepackage{amsthm,amsmath}
%\RequirePackage{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
%\RequirePackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
\usepackage{color}
% \usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\def\includegraphic{}
\def\includegraphics{}

%%% Put your definitions there:
\startlocaldefs
\endlocaldefs

%%% Begin ...
\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Research}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{A game-theory modeling approach to utility and strength of interactions dynamics in biomedical research social networks}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author[
   addressref={aff1,aff2},                   % id's of addresses, e.g. {aff1,aff2}
   corref={aff1},                       % id of corresponding address, if any
   noteref={n1},                        % id's of article notes, if any
   email={jmario.siqueiros@iimas.unam.mx}   % email address
]{\inits{JMSG}\fnm{J. Mario} \snm{Siqueiros-Garc\'ia}}
\author[
   addressref={aff2},
   email={rgarcia@ie.unam.mx}
]{\inits{RGH}\fnm{Rodrigo} \snm{Garc\'ia-Herrera}}
\author[
   addressref={aff3},
   email={ehernandez@inmegen.mx}
]{\inits{EHL}\fnm{Enrique} \snm{Hern\'andez-Lemus}}
\author[
   addressref={aff3},
   email={sergio.alcala@ciencias.unam.mx}
]{\inits{SAC}\fnm{Sergio} \snm{Alcal\'a-Corona}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\address[id=aff1]{%                           % unique id
  \orgname{IIMAS, Universidad Nacional Aut\'onoma de M\'exico}, % university, etc
  \street{Circuito Escolar 3000, Cd. Universitaria},                     %
  \postcode{04510}                                % post or zip code
  \city{Mexico City},                              % city
  \cny{Mexico}                                    % country
}
\address[id=aff2]{%
  \orgname{LANCIS-IE, Universidad Nacional Aut\'onoma de M\'exico},
  \street{Ciudad Universitaria},
  \postcode{04510}
  \city{Mexico City},
  \cny{Mexico}
}
\address[id=aff3]{%
  \orgname{Instituto Nacional de Medicina Gen\'omica},
  \street{Periferico Sur 4809},
  \postcode{14610}
  \city{Mexico City},
  \cny{Mexico}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter short notes here                   %%
%%                                          %%
%% Short notes will be after addresses      %%
%% on first page.                           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{artnotes}
%\note{Sample of title note}     % note to the article
\note[id=n1]{Equal contributor} % note, connected to author
\end{artnotes}

\end{fmbox}% comment this for two column layout

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}

\begin{abstract} % abstract
%\parttitle{First part title} %if any
Collaboration has become a cornerstone in biomedical research today.
In contrast to physics which has a long history and experience in
collaborative projects, biology is only recently becoming an evermore
collaborative discipline.
In this article we explore the effect of a collaboration network on the distribution
of players having access to certain amount of resources from other players in the
network and the distribution of the strength of connections among
them. Particularly, we implemented two games played simultaneously:
one for maximazing individual utility based on the iterated Prisoner's
Dilemma; the other, a coordination game for maximazing the connection strength
between players. We are interested in how they affect each other in 
the context of a network of scientific collaboration under the idea that while
researchers are interested in maximazing their utilities, they also know that
it is important to invest in building collaborative relationships.
We tested our simulation on a biomedical research community network of M\'exico and
compared the results with an Erd\"{o}s-Reny\'i, a Watts-Strogatz
small-world and Barab\'asi-Albert topologies. Different topologies
display different global utility and global strength of interaction
distributions. Moreover, the distribution of utility and strength of interaction
in the researchers network is similar to that of 
Barab\'asi-Albert and Watts-Strogatz topologies, respectively. We
believe that utility distribution in the researchers network
suggests that there are socio-cultural mechanisms governing the
network that produce an asymmetric distribution of resources. The
high distribution of strong connections might reflect some sort of
subordination among researchers by which they are morally obliged to
cooperate by the same socio-cultural mechanisms. Finally, the range around
the threshold that regulates the decision to cooperate or defect
according to the agent's historical balance between utility and
strength of collaborative relationships and carrying capacity of the
system is small, suggesting that there is a region in which a phase
transition takes place from a population of cooperators to a
population of defectors. Simulations like ours may help to develop
science policies to promote fair distribution of resources.

%\parttitle{Second part title} %if any
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The keywords begin here                  %%
%%                                          %%
%% Put each keyword in separate \kwd{}.     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{keyword}
\kwd{Collaboration networks}
\kwd{Social Networks Analysis}
\kwd{Biomedicine}
\kwd{Game theory}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
%
%\end{fmbox}% uncomment this for twcolumn layout

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Main Body begins here                %%
%%                                          %%
%% Please refer to the instructions for     %%
%% authors on:                              %%
%% http://www.biomedcentral.com/info/authors%%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%% See the Results and Discussion section   %%
%% for details on how to create sub-sections%%
%%                                          %%
%% use \cite{...} to cite references        %%
%%  \cite{koon} and                         %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
% <put your article body there>

%%%%%%%%%%%%%%%%
%% Background %%
%%

\section*{Introduction}

Collaboration has become a cornerstone in biomedical research today.
In contrast to physics which has a long history and experience in
collaborative projects, biology is only recently becoming an evermore
collaborative discipline\cite{VermeulenPenders:2013}. Biology has an
interesting record in such matters because scientific collaboration
means something different to different branches of biology: molecular
biology has traditionally been a research activity of small
laboratories\cite{KnorrCetina:1999,Strasser:2006}, whereas in natural
history there has been data and samples exchange since the $XVII^{th}$
century\cite{MullerWille-etal:2012,Strasser:2012}. Despite the differences in
culture and practices, the Human Genome Project made collaboration a
central feature of biology.\\

@@

The analysis of cooperation in scientific research has been the subject of a
number of studies
\cite{VermeulenPenders:2013,Newman:2001,Newman:2004,Elango-etal:2012,HernandezLemus:2013,Strasser:2006,Strasser:2012}. This
is not surprising since cooperation and competition are quite important in
today's academic success. How does collaboration happen within a competitive
academic environment and what kind of payoff is present in these settings were
questions considered recently by Wardil and Hauert \cite{Wardil-etal:2015} in the
context of cooperation in multi-authored publications. Also, the role of game
theory over complex scientific information and collaboration networks has
attracted attention, mainly focusing on how long-term strategies may shape
different scenarios for Nash equilibria \cite{Hanauske-etal:2012}. Prisoner's Dilemma
has been used in the study of impact factor and collaboration
\cite{Hara-etal:2003,LiebermanNowak:2005}. \\

Even with all these research efforts, cooperation in the context of
scientific collaboration is still loosely defined and the long term
dynamics of academic cooperation (and its consequences) are yet to be
fully elucidated. Furthermore, to our current knowledge, there has
been no use of game theory and complex network analysis for
understanding how the topology of scientific collaboration networks
affects access to resources among individuals present in the network
\footnote{For an account of scientific collaboration and definitions,
  please refer to \cite{Sonnenwald:2007}.}. Our work aims to contribute
to our current understanding on the matter, specially when agents have
to maximize their access to resources while taking care of their collaboration links.\\

 In this article we explore the network effect on the distribution
   of players having access to certain amount of resources shared by other players in the
   network and the distribution of the strength of interactions among
   them. Particularly, we implemented two games played simultaneously: 
   one for maximazing individual utility based on the iterated Prisoner's
   Dilemma; the other, a coordination game for maximazing the connection strength between
   players. We are interested in how they affect each other in 
   the context of a network of scientific collaboration under the idea that while
   researchers are interested in maximazing their utilities, they also know that
   it is important to invest in building collaborative relationships. These two
   behaviors are explored in a biomedical research community of M\'exico.\\

In the context of our model, utility represents
  access to resources shared by others.
  The value of the utility function for a player is the sum of the payoffs of playing
  with its neighbors. The opossing force comes from the other concurrent game: players
  trying to maximize the strength of their interactions with other players. In the coordinating game
  the best strategy is to adopt the same strategy as the other player, as it pays the most,
  regardless of cooperation or defection in the utility game. When
  both cooperate the interaction gets a positive payoff, when both defect, the
  interaction doesn't get affected; but if they anti-coordinate, then the
  interaction looses. Finally, cooperation is a central feature of scientific work.
  For our biomedical network, cooperation can be thought of
  as sharing resources such as time, students, equipment, even money. Examples of defection
  to a cooperator are ghost authorship or
  prestige authorship.\\

The manuscript is structured in four sections. First we describe
\textit{FOSISS}, the main program for grants destined to biomedical applied
research in M\'exico. This is the source of the
database from which we created the researchers collaboration network. Next we
describe our model
and the different network topologies on which we explored it.
We then present our results and discuss them. In the last section
we draw some final remarks and conclusions.

\section*{Biomedical research: CONACyT and FOSISS}

@@

\emph{FOSISS} or Sectoral Fund for Health and Social Security Research
(\emph{Fondo Sectorial en Investigaci\'on en Salud y Seguridad Social}) is one of
such funds. FOSISS is constituted by CONACyT, SSA, IMSS and ISSSTE,\footnote{SSA
  is the acronym for Secretariat of Health \emph{Secretar\'ia de Salud}; IMSS is
  the acronym for Social Security Mexican Institute (\emph{Instituto Mexicano
    del Seguro Social}); ISSSTE stands for Institute for Social Security and
  Services for State Workers (\emph{Instituto de Seguridad y Servicios Sociales
    de los Trabajadores del Estado})} all of them being the major public health
providers and research institutions in the country. Every year CONACyT opens a
call for funds limited to a set of health research areas previously defined by a
group of experts. Most applicants are public universities and research
institutions, but eligibility is open to public and private health research sectors.
From 2002 through 2013, there were 91 institutions funded that
comprised 4988 researchers.\\

@@

\section*{Methodology}

%% In this work we developed a model in order to have a better understanding of the
%% distribution of utility and trust in real networks of biomedical research
%% collaboration under certain social constraints  or topological properties.\\

Our model is based on the iterated version of the Prisoner's Dilemma (PD) and a
coordination game instantiated on
networks. Implementing games on networks is not new and it's an active area of
research aimed to understand the evolution of cooperation in networks populated
by selfish agents \cite{SzaboFath:2007,Nowak-etal:92,OshtukiNowak:2006,Santos-etal:2005,Santos-etal:2006}. In many network models
on which some of game theory games are simulated, agents' decision to cooperate
or defect depend on a specific strategy, such as the well known
\textit{tit-for-tat} \cite{Axelrod:2006,Nowak-etal:2011}. In some other cases, agents
can modify the weight of the interactions with their neighbors
\cite{Santos-etal:2006}. From a different perspective others have explored the effect
of different topologies on the emergence of cooperation
\cite{Santos-etal:2005,HauertDoebeli:2004}. In our model, an agent's decision to
cooperate or defect depends on a balance between
utilities and the current strength of its collaboration relationships. Such
balance reflects the overall success or failure of its strategies. We
study the behavior of the system under different topologies, including a
real-world network.\\

@@

\subsection*{Network initialization and agent state update}

All networks are initialized equally. The number of nodes for every network is 4122, the same as in the FOSISS network.
The same utility is given to every agent and all edges are asigned the same weight. In the case of the FOSISS
network, edge weight is given by the number of collaborations among researchers, utility remains the same for all nodes as in
the other networks.\\

The probability for an agent to cooperate or defect depends on a number ($\eta$)
that referes to a historical balance between average utility and the average
strength of the connections with its neigbors. This is so because we assume that
whatever the result in utility or strength of connection, as long as one of them
increases, the player will be confident in the strategy followed so
far. In every simulation step, each player will play both games
  with every neighbor, but set its decision to cooperate or defect beforehand by
  assesing the overall situation of its relationships. This may sound odd, since
  one may think that a person decides for each relationship separately. However
  in this way players are optimizing access to resources and the strength of
  interactions to their neighbors at the same time, since their state is set by
  considering both together.\\

@@

\subsection*{Erd\"{o}s-Reny\'i}
